---
title: "R DAPC"
author: "Dr. Wilson Frantine-Silva"
date: '2022-06-22'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DAPC analysis with ggplot2

This is a simple example of how to run a simple DAPC analysis with SSR or SNIP data.

For details about DAPC, see the adegenet package documentation and Joumbart et al., (2010). 

DAPC analysis is very useful for the exploratory analysis of genetic data. It does not rely on Hardy-Weinberg equilibrium assumptions or other genetic models our data often breaks.

There are roughly two ways to run a DAPC analysis. But before get deeper into it, lets get the libraries we are going to use:


```{r}
library("adegenet") # to run DAPC
library("tidyverse") # to use pipe %>% and other functions
library("ggplot2") # to get better plots
library("scico") # to get good scientific color pallets
library("patchwork") # to make good compositions
```

## Importing data 

The adegenet package accepts several types of data, but the genepop is one of most popular for sure.    
 > ! Make sure your file has .gen as extension.

```{r}
file <- "~/genepop.gen"
```

After that, lets read it as an genepop file with `read.genepop()` function, and create a summary of it.
 > ! make sure to have `ncode` equal to the number of character used to code each alellel!

```{r}
data <- adegenet::read.genepop(file, ncode = 3)
summary(data)
```


```{r}
```


